---
author: kevin peng
title: "Lab 19: Inference for Several Means"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.height = 5)
knitr::opts_chunk$set(fig.width = 8.5)
knitr::opts_chunk$set(out.width = "100%")
knitr::opts_chunk$set(dpi = 300)

library(readr)
library(ggplot2)
library(dplyr)
library(viridis)
library(forcats)
library(smodels)
theme_set(theme_minimal())
```

## Instructions

Below you will find several empty R code scripts and answer prompts. Your task
is to fill in the required code snippets and answer the corresponding
questions.

## Burrito synergy in SoCal

Today we will look at a dataset of reviews of burritos in southern
California. This is a data set collected by a group of college friends who
live in the greater San Diego area.

```{r}
burrito <- read_csv("https://statsmaths.github.io/stat_data/burrito.csv")
```

The available variables in the data are:

- location: the name of the restaurant
- cost: total cost for the burrito
- yelp: the average Yelp review score
- google: the average Google review score
- chips_included: equals 1 if chips are included with the burrito
- hunger: score from 1 to 5 on how much the burrito filled up the reviewer
- tortilla: rating from reviewer; 1 (terrible) to 5 (awesome)
- temp: rating from reviewer; 1 (terrible) to 5 (awesome)
- meat: rating from reviewer; 1 (terrible) to 5 (awesome)
- fillings: rating from reviewer; 1 (terrible) to 5 (awesome)
- meat_filling: rating from reviewer; 1 (terrible) to 5 (awesome)
- uniformity: rating from reviewer; 1 (terrible) to 5 (awesome)
- salsa: rating from reviewer; 1 (terrible) to 5 (awesome)
- synergy: rating from reviewer; 1 (terrible) to 5 (awesome)
- wrap: rating from reviewer; 1 (terrible) to 5 (awesome)
- overall: rating from reviewer; 1 (terrible) to 5 (awesome)
- lat: latitude (in degrees North) of the taco restaurant
- lon: longitude (in degrees East) of the taco restaurant

Fit a linear regression predicting the average yelp score of locations in this
dataset and look at the model using `reg_table`:

```{r}
model <- lm_basic(yelp ~ 1, data = burrito)
reg_table(model, 0.95)
```


The average Yelp score of all restaurants in this area is just 3.7; looking
at the confidence interval from the model does it seem likely that burrito
locations have a higher Yelp rating than locations in general, or is this
just noise in the data?

**Answer**: Yes. Burrito shops in this location have an estimated average of 3.93 as the average Yelp review. With the 95% confidence interval, we can see that we are 95% confident that the true mean lies between 3.808 and 4.052, which is larger than 3.7.

The regression model confidence interval requires that the data are sampling
independently from the larger population. Why might this dataset not hold to
this assumption?

**Answer**: The data is only taken from San Diego, CA, and even so, it was collected by a group of friends from there with no sign of being indepentent. 

Now fit a linear regression that predicts the yelp score as a function of
whether the burrito includes chips.

```{r}
model1 <- lm_basic(yelp ~ 1 + chips_included, data = burrito)
reg_table(model1, 0.95)

```

Describe what the slope coefficient in the previous model really means
in terms of the data:

**Answer**: The slope coefficient is the 0.05699 under the estimate column. The meaning is that you need to add 0.05699 to get the mean of yelp reviews of places where chips were included.

Is the slope coefficient statistically significant (compared to zero)?

**Answer**: No. 

Describe this in terms of your answers above.

**Answer**: We would not be surprised if the chips made the yelp scores -0.2 lower or 0.3 higher and that does not result in significance because those numbers average to near 0.

The word *synergy* can be defined as:

> the interaction or cooperation of two or more organizations, substances,
> or other agents to produce a combined effect greater than the sum of
> their separate effects.

This is not something I would have thought of when considering tacos, but
let's see how synergy effected the scores of both the reviewers as well as
Google and Yelp users in general.

First, I'll add a variable called high_synergy defined as whether the synergy
score is greater than 3.5.

```{r}
burrito$high_synergy <- (burrito$synergy > 3.5)
```

Fit a regression predicting the Yelp score as a function of the
variable `high_synergy`.

```{r}
model2 <- lm_basic(yelp ~ 1 + high_synergy, data = burrito)
reg_table(model2, 0.95)

```

Is there strong evidence that Yelp scores are higher for high synergy burrito
restaurants?

**Answer**: Yes. We would not be surprised with a variance from the mean to be between -0.00927 and 0.485 and more often than not that number would result in an overall increase in score due to the range being higher. 

Now fit a regression predicting the overall score as a function of the
variable high_synergy.

```{r}
model3 <- lm_basic(overall ~ 1 + high_synergy, data = burrito)
reg_table(model3, 0.95)

```

Is there evidence that high synergy burrito locations have a higher overall
score than low synergy burrito locations?

**Answer**: Yes. We are 95% confident that there will be an increase between (0.8703-1.465) in its overall score if there is a high synergy.

Using the model from the previous question, what does the model predict
will be the overall score of a location with low synergy?

**Answer**: The model predicts that the true mean of a store without synergy will be between 2.4583 and 2.934, with a estimated mean of 2.6963. 

What does the model predict will be the overall score of a location with
high synergy?

**Answer**: The model predicts the overall score of a location with high synergy will be 2.6963 + 1.1678 = 3.8641.

Finally, fit a regression model that predicts the overall score as a function
of the variable `synergy` coded as a factor.

```{r}
model4 <- lm_basic(overall ~ 1 + factor(synergy), data = burrito)
reg_table(model4, 0.95)

```

What does the model predict the overall score for a restaurant with a
synergy score of 3 will be?

**Answer**: The model predicts the overall score of restuaurant with synergy 3 is 2.50 + 0.6917 = 3.1817.





