---
title: "Lab 21: Sixteen Personality Factors I"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.height = 5)
knitr::opts_chunk$set(fig.width = 8.5)
knitr::opts_chunk$set(out.width = "100%")
knitr::opts_chunk$set(dpi = 300)

library(readr)
library(ggplot2)
library(dplyr)
library(viridis)
library(forcats)
library(smodels)
theme_set(theme_minimal())
```

## Instructions

Below you will find several empty R code scripts and answer prompts. Your task
is to fill in the required code snippets and answer the corresponding
questions.

## Sixteen Personality Factors: Pick Your Trait

Over the next two classes we are going to be looking at survey response
data from the Sixteen Personality Factor Questionnaire in order to practice
our skills at statistical inference:

```{r}
pf <- read_csv("https://statsmaths.github.io/stat_data/cattell_16pf.csv")
```

The dataset use the following fields:

- age: respondent's age in years
- gender: respondent's self-selected gender
- country: two letter IATA code for the respondent's IP
- elapsed: time taken to complete quiz in seconds
- warmth: personality score from 1-20
- reasoning: personality score from 1-20
- emotional_stability: personality score from 1-20
- dominance: personality score from 1-20
- liveliness: personality score from 1-20
- rule_consciousness: personality score from 1-20
- social_boldness: personality score from 1-20
- sensitivity: personality score from 1-20
- vigilance: personality score from 1-20
- abstractedness: personality score from 1-20
- privateness: personality score from 1-20
- apprehension: personality score from 1-20
- openness_to_change: personality score from 1-20
- self_reliance: personality score from 1-20
- perfectionism: personality score from 1-20
- tension: personality score from 1-20
- baseline: average score across all 16 personality traits

*To start with, select a particular trait that you will use for the first
bank of questions. I suggest picking something that popped out when you took
the test. You can pick any of the 16 other than sensitivity.*

Produce a bar plot of the personality scores for your trait.

```{r}
ggplot(data = pf, aes(rule_consciousness)) +
         geom_bar()

```

Describe the distribution. Why does a bar plot work here even though the
variable is numeric?

**Answer**: It works because there will be frequencies of values 1-20 and you can just plot each one.

Produce a confidence interval for the mean of your trait.

```{r}
model <- lm_basic(rule_consciousness ~ 1, data = pf)
reg_table(model, level = 0.95)
```

Now, produce a dataset that consists only of responses from the country of
Hungary. The 2-letter country code for Hungary is "HU". Produce a confidence
interval for your trait on the Hungarian dataset.

```{r}
hungary <- filter(pf, country == "HU")
model <- lm_basic(rule_consciousness ~ 1, data = hungary)
reg_table(model, level = 0.95)
```

What do you notice about this confidence interval compared to the original
one? Can you explain why this is the case?

**Answer**: What I noticed was that the intercept of rule_consciousness in Hungary was ~1.4 units less than the whole dataset. A reason this may be the case is that Hungary may not be as fully developed as a country compared to the USA, where much of the whole dataset is from. 

```{r}
model <- lm_basic(rule_consciousness ~ 1 + gender, data = hungary)
reg_table(model, level = 0.95)
```

Find a 95% confidence interval for the difference between the average
male and female value for your personality trait on the Hungarian data.

**Answer**: 
Female = (11.4103, 12.590)
Male = (11.4103 - 1.0424, 12.590 + 0.781) -->
Male = (10.3679, 13.371)



Is there a statistically significant difference between men and women's
average score? If so, what direction is this difference? Does this challenge
or confirm traditional gender stereotypes (note: not all personality traits
have one)?

**Answer**: No. The R-squared value is close to 0, which means it is not very significant. 

Construct a new dataset that only has ages from 30-49 (from the original `pf`,
not just the Hungarian subset). Hint: you can use the filter function twice.

```{r}
over30 <- filter(pf, age >= 30)
over30Under50 <- filter(over30, age <= 49)
over30Under50$fourties <- (over30Under50$age >= 40)
```

On the dataset of people ages 30 to 49, create a variable called `fourties` if
age is greater than or equal to 40.


Fit a regression on with your personality trait as a response to test for the
difference between its mean for people in their 30s versus people in their
40s. Compute a confidence interval for this difference.

```{r}
model <- lm_basic(rule_consciousness ~ 1 + fourties, data = over30Under50)
reg_table(model, level = 0.95)

```

Is there a statistically significant difference between 30s and 40s average
scores? If so, what direction is this difference? Does this challenge or
confirm traditional age stereotypes (note: not all personality traits
have one)?

**Answer**: The multiple R-squared's value is very close to 0, and as a result, there is no statistical significant difference between 30s and 40s. 

Draw a bar plot of the variable `country` over the whole dataset. Take
note of the number of countries with a very small number of responses.

```{r}
ggplot(pf, aes(country)) + 
  geom_bar(aes(fct_lump(country, n = 5)))
```

Fitting a model with all of these countries is possible but not very
useful. When we only have a few responses from some places it makes those
few regions with a lot of data harding to identify and analyze. Fortunately
there is a solution in the function `fct_lump`.

Fit a regression model on the entire dataset by the country variable lumped
into 5 categories.

```{r}
model <- lm_basic(rule_consciousness ~ 1 + fct_lump(country, n = 5), data = pf)
reg_table(model, level = 0.95)

```

Using the previous model, which country (not including "Other") has the
highest score for your trait? Which has the lowest?

**Answer**: 
Rule_consciousness
Highest score: IN
Lowest score: GB

## Fitting Sensitivity Scores

Fit a model that predicts sensitivity as a function of the baseline
score.

```{r}
model <- lm_basic(sensitivity ~ 1 + baseline, data = pf)
reg_table(model, level = 0.95)

```

Is the slope statistically significantly different from 1? Why is this an
interesting question in the context of the data?

**Answer**: Yes. The R-Squared value is closer to 1 than not. In the context of this data, it shows how sensitivity positively affects the baseline score.

Fit a regression model that uses both gender and the baseline score to
predict your trait.

```{r}
model <- lm_basic(sensitivity ~ 1 + gender + baseline, data = pf)
reg_table(model, level = 0.95)
```

Describe the slope for the gender term in words:

**Answer**: If the gender is male, the sensitivity score is -0.5495 lower than females. 

Add predictions from the previous model to `pf`.

```{r}
model <- lm_basic(sensitivity ~ 1 + gender + baseline, data = pf)
pf <- add_prediction(pf, model)
reg_table(model, level = 0.95)



```

Plot the baseline score as a function of `model_pred`, coloring the points
base on the gender variable.

```{r}
ggplot(pf, aes(baseline, sensitivity)) +
  geom_point() +
  geom_point(aes(y = model_pred, color = gender)) +
  geom_line(aes(y = model_pred, color = gender))

```

What do the predicted values look like?

**Answer**: Looks linear. As sensitivity increases, the baseline generally increases. Also, females are higher

