---
title: "Kevin Peng vs. Ryan Namgung, a Basketball Analysis"
author: "Kevin Peng"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(readr)
library(dplyr)
library(plyr)

```


## Overview

My roommate and I have a slight rivalry on the basketball court. In order to determine
who is the best, we decided to have a 1v1 match and record stats for each possession.
We played the game as first to 50 points, counting shot values as 2-pointers or 3-pointers.
Statistics recorded in this project included variables such as shot type, 
shot number, defense level, and more. The game was filmed and data was extracted from it. 
When collected, the data was then seperated into two serperate datasets, one for me and one for Ryan, with the same variables
for clarity.
Ultimately, this research is not only used to determine who is better at basketball. It 
also provides the opportunity to learn our strengths/weaknesses and 
analyze tendencies/references on the basketball court. This data will 
provide us with parts of the game that each of us need to improve on. 

## Data dictionary

Here is a data dictionary for the kevin_basketball and ryan_basketball datasets:

### numeric variables
- possession: a numeric variable that determines which possession of the game either Ryan and I are at, independently.
- shot_value: a numeric variable that describes the value of a given shot, a two or a three pointer.
- field_goals_attempted: a numeric variable that describes how many total shots each player has taken.
- field_goals_made: a numeric variable that describes how many total shots each player has made. 
- three_pointers_attempted: a numeric variable that describes how many threes each player has taken.
- three_pointers_made: a numeric variable that describes how many threes each player has made.
- score: a numeric variable that keeps track of the score as the game progresses, independent for each player. 

### categorical variables
- player_name: a categorical variable that determines who has possession of the ball at a given time. The format will be player's initials, "K" for Kevin, and "R" for Ryan. 
- shot_type: a categorical variable that describes that type of shot taken by a player. Possible options are:
    - "layup"
    - "jumpshot"
    - "fadeaway" (jumpshot with a backwards jump)
- shot_contestion: a categorical variable that describes the level of defense played during the point. Possible levels are "low, medium, high".
- make_miss: a categorical variable that describes if a shot was made of missed. 



The datasets can be read in as follows:
```{r message=FALSE}
kevin_basketball <- read_csv("https://raw.githubusercontent.com/URstats/stat209-s18-kevpeng/master/kevin_data.csv")
ryan_basketball <- read.csv("https://raw.githubusercontent.com/URstats/stat209-s18-kevpeng/master/ryan_data.csv")
```


```{r echo=FALSE}
(kevin_basketball)
(ryan_basketball)
```

## Exploratory Analysis


### Score Progression

Throughout the game, I knew my strengths were shooting three pointers. As a result, I did my best to shoot as many of them as I can. The following plot shows what value shot I took each possession, along with a line to show what point is victory.  

```{r, echo=FALSE}
ggplot(kevin_basketball, aes(possession, score)) +
  geom_point(aes(color=factor(shot_value))) + 
  ggtitle("Kevin Shot Values") +
  geom_hline(yintercept = 50, col = "red")
```


The following plot shows what value shots Ryan took each possession.
```{r, echo=FALSE}
ggplot(ryan_basketball, aes(possession, score)) +
  geom_point(aes(color=factor(shot_value))) + 
  ggtitle("Ryan Shot Values") +
  geom_hline(yintercept = 50, col = "red")

```

From the data, it can be observed that I win by reaching 50 points first. In 
addition, while comparing these plots, the score line on mine has steeper slopes
than Ryan's. This is attributed to scoring streaks, where I score
multiple buckets in a row. Ryan's plot has lower shopes because he would 
not go on scoring streaks, frequently missing after making one. 

### 3-Pointers

I decided to dig deeper into the three pointer data. 
I filtered the data into a new dataset that only contained shots
that were 3-pointers. I then plotted it with the shot make/made as 
an aestetic. I took the data further by filtering it based 
on shot_contestion, or defense level and examined shooting 
percentages based on different defense. 

```{r, echo=FALSE}
three_pointers_kevin <- filter(kevin_basketball, shot_value == 3)
ggplot(three_pointers_kevin, aes(possession, score)) +
  geom_point(aes(shape=make_miss, color=shot_contestion), size = 1.5) + 
  ggtitle("Kevin 3-Point Shot")

# further filtering
three_kevin_low  <- filter(three_pointers_kevin, shot_contestion == "low")
three_kevin_med  <- filter(three_pointers_kevin, shot_contestion == "med")
three_kevin_high <- filter(three_pointers_kevin, shot_contestion == "high")

# low defense percent
three_low_kev_percent = sum(three_kevin_low$make_miss == "make") / length(three_kevin_low$make_miss)
print("kevin three pointer %, low defense")
print(three_low_kev_percent)

# med defense percent
three_med_kev_percent = sum(three_kevin_med$make_miss == "make") / length(three_kevin_med$make_miss)
print("kevin three pointer %, medium defense")
print(three_med_kev_percent)

# high defense percent
three_high_kev_percent = sum(three_kevin_high$make_miss == "make") / length(three_kevin_high$make_miss)
print("kevin three pointer %, high defense")
print(three_high_kev_percent)

# overall three shooting %
three_percentage_kev = (max(three_pointers_kevin$three_pointers_made) / 
    max(three_pointers_kevin$three_pointers_attempted))
print("kevin three pointer %, overall")
print(three_percentage_kev)


```


```{r, echo=FALSE}
three_pointers_ryan <- filter(ryan_basketball, shot_value == 3)
ggplot(three_pointers_ryan, aes(possession, score)) +
  geom_point(aes(shape=make_miss, color=shot_contestion), size = 2) + 
  ggtitle("Ryan 3-Point Shot")
three_percentage_ryan =   
  (max(three_pointers_ryan$three_pointers_made) / 
    max(three_pointers_ryan$three_pointers_attempted))
print("ryan three pointer %, overall")
print(three_percentage_ryan)
```

From the data, there is a clear winner, at least in
3-point shooting percentage. Overall, I shot 36.84%, while Ryan shot 6.25%. 

I looked at the defense levels ability on my shooting percentages. 
While the defense level is low, my shooting percentage is the highest, which
is to be expected, at 40%. What was surprising was how having medium defense 
was more detrimental to my shot than high defense, with a 25% chance for medium 
defense, and a 36% for high defense. Having a high level of defense only dropped 
my shot by ~4% vs. having low defense. This data suggests that playing the right
level of defense is required to decrease my shooting percentage.

For Ryan, all of his 3-point attempts had low defense, so it was not filtered into detail.

### 2-Pointers

I filtered data into only 2-pointers as well. Once filtered to only 2 pointers, 
I wanted to see how we both fare shooting from inside the 3-point line. 

```{r, echo=FALSE}
two_pointers_kevin <- filter(kevin_basketball, shot_value == 2)
ggplot(two_pointers_kevin, aes(possession, score)) +
  geom_point(aes(shape=make_miss, color=shot_type), size = 3) + 
  ggtitle("Kevin 2-Point Shot")

# overall two shooting %
two_percentage_kev = 6/23
print("kevin two pointer %, overall")
print(two_percentage_kev)

```

```{r, echo=FALSE}
two_pointers_ryan <- filter(ryan_basketball, shot_value == 2)
ggplot(two_pointers_ryan, aes(possession, score)) +
  geom_point(aes(shape=make_miss, color=shot_type), size = 2) + 
  ggtitle("Ryan 2-Point Shot")

# overall two shooting %
two_percentage_ryan = 20/42

print("ryan two pointer %, overall")
print(two_percentage_ryan)


```

Based on the data, we are able to see where Ryan gets the majority of his
points from. Compared to his 3-point percentage, his 2-point percentage is 
excellent. His 47.6% is better than my 26% from 2-point range. Additionally,
Ryan shot approximately twice as many 2-pointers than I did, 42 to 23. This
data shows that Ryan is more comfortable driving in and shooting from close range,
and he has the advantage doing so. 


## Conclusion

From this project, Ryan and I were able to determine a winner. I ultimately won out,
with a final score of 51-43. In addition, I was able to calculate both of our shooting
percentages, inside and outside the three point line. I learned that my tendency and preference is to shoot 
from long distance, attempting 38 3-pointers to 23 2-pointers, with only 7 layup attempts. 
Ryan's preference was shooting from close range, attempting 17 3-pointers to 42 2-pointers, with 20 layup
attemps. Ryan's results may be skewed because he attempted a large number of 3-pointers, missing the last 16 in a row. 
I recommended that he improve shooting from outside, or shoot more from the inside, to his advantage. 
I was also able to determine that I must work on making shots when closer to the basket. 

Other variables that I would like to add to this project pertain to misses. 
I would include which direction the ball missed: left, right, far, or short. 
From an analysis I would be able to determine how to improve based on which direction was
most frequent. For example, if the ball misses left, shoot more right. 

Ultimately, basketball is a game of numbers and whoever gets more numbers, will be the winner. 

## References
### Special thanks to:
  - Joseph Mugisha for lending me his camera to record the game.
  - David Qin for recording the battle.
  - Ryan Namgung for agreeing to lose.



