---
title: "Kevin's 3 Pointers"
author: "Kevin Peng"
output: html_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(out.width = "100%")
knitr::opts_chunk$set(dpi = 300)

library(readr)
library(ggplot2)
library(dplyr)
library(ggmap)
library(viridis)
library(forcats)
library(smodels)
```

## Introduction

Similar to my first project, I decided to analyze my own basketball game. 
Instead of playing a 1v1 game versus someone, I decided to record shots from 
around the 3-point line and determine shooting patterns. The three hypothesis 
that I explore are:


1.) For any given, unguarded, practice shot, I have >= 50% chance of making it. 

2.) After making a shot, the next shot will have a higher chance of going in.

3.) I am more likely to go on a make streak from the left side of the court compared to 
    the right side.

```{r, include=FALSE}
three_data <- read_csv("https://raw.githubusercontent.com/URstats/stat209-s18-kevpeng/master/project-c.csv")
```

## Methods

In order to collect this data, I asked a friend to sit on the side and
record me taking 100 shots, 20 from 5 different points around the three point
line: The corners, the wings, and the top of the key, looking towards the baseline 
(Figure 1). With each shot, there were variables considered, which are in the data dictionary below. 

The inferencial models used for linear regression were: 

\t 1.) shots_made compared to the shot_number

\t 2.) shot_made_streak compared to the shot_number 

The model used for inference of the mean: 

\t 1.) shot_made_streak compared to location of the shot. 


#### Data dictionary for the dataset "three_data":
- location: location on the court where a shot was taken. Options can be seen around the three point line of Figure 1.
- shot_number: a numeric variable 1-100 describing which shot is being taken.
- shots_made: a numeric variable counter that describes how many shots were made
- shot_made_streak: a numeric variable that acts as a counter keeping track of consecutive shots made.
- made_miss_streak: a categorical variable describing the streak type that is happening at a given shot_number. Options include: miss_X, and make_X, where X is an positive integer of the streak.

[Figure 1. Shot locations](https://www.coachesclipboard.net/images/xHalfCourtTerms.gif.pagespeed.ic.f_tQMgajQe.png)




## Results

### Linear Regression

```{r, include=TRUE}
model <- lm_basic(shots_made ~ 1 + shot_number, data = three_data)
reg_table(model, level = 0.95)
qplot(shot_number, shots_made, data = three_data) + geom_smooth(method="lm")
```

This plot shows a line y = ax + b, where y is the number of shots made, x is the number of shots taken and a/b are the intercept and slope from the linear regression. The formula and plot both show a positive relationship between the variables, which is expected, and a line that predicts, if I take x-amount of shots, I will make y of them. 


```{r}
model <- lm_basic(shot_made_streak ~ 1 + shot_number, data = three_data)
reg_table(model, level = 0.95)
qplot(shot_number, shot_made_streak,data = three_data) + geom_smooth(method="lm")


```

This plot shows a line y = ax + b, where y is the shooting streak, x is the number of shots taken and a/b are the intercept and slope from the linear regression. The intercept of the line is 1.12, and the slope when considering shot_number is -0.0058. Plugging in a number x would determine what the average streak I go on is. 

### Inference for the mean

```{r}
model <- lm_basic(shot_made_streak ~ 1 + location, data = three_data)
reg_table(model, level = 0.95)
ggplot(three_data, aes(location, shot_made_streak)) +
  geom_violin()
```

The violin plot shows the relationship between locations and average shot streak.
The regression table shows how each location affects the average shot streak, each named
based off Figure 1. 

## Conclusions

#### Linear regression test
The relationship between shots_made and shot_number show a positive relationship, and more importantly, above 0.50 for the slope when considering shots_made. When I plug in a number 100, for the shot_number, the linear regression predicts a number of shots made as ~51.3 made. Combined with the multiple R-squared statistic being close to 1 and the intercepts and slope not including 0, I am confident with the result and that there is at least a 50% chance that I make a wide open 3-point shot, and can confirm the first hypothesis.   

The relationship between shots_made and shots_made_streak show a negative relationship. 
We are 95% confident that the true mean of the shots_made_streak will be 1.213, between 0.7518 and 1.675. The positive numbers of the confidence interval show that this number is statistically significant. As a result, we can confirm the second hypothesis that if a shot is made, the probability of the next shot is higher. 

The slope on the other hand, was between -0.013743 and 0.002. Despite the upper bound being barely above 0, I would not consider the shot number as being statistically significant, as in addition, the multiple R-squared statistic is close to 0. As a result, we can also conclude that shot number does not affect the probability of making consecutive buckets. 

Despite rejecting the second hypothesis, the plot for it showed an interesting trend. As I shot more, I was going on less streaks. I believe this may have to do with fatigue as I shot more shots.

#### Inference for the mean

The left_corner (intercept) was the only location that had statistical significance, 
with both parts of the confidence interval being positive. For that location, 
I am 95% confident that the true mean of the shot_make_streak was 1.91667, between 1.3235 and 
2.510. However, because the other's datas were not significant, as their confidence interval all contained 0, and the R-squared statistic was too close to 0, we reject the third hypothesis that location affected the average shot_make_streaks. 


## References

1. https://www.coachesclipboard.net/images/xHalfCourtTerms.gif.pagespeed.ic.f_tQMgajQe.png


